---
title: "Business Analytics"
subtitle: "Advanced Data Visualizations"
author: "Ayush Patel and Jayati Sharma"
date: today
date-format: " DD MMMM, YYYY"
embed-resources: true
format: 
  revealjs:
    embed-resources: true
    slide-number: c/t
    width: 1400
    theme: [serif, theme.scss]
---

## Pre-requisite {.scrollable}

::: incremental
You already....

-   Know basic and advanced data wrangling functions in R
-   Know basics of data visualization in R
-   Can write functions in R
:::

## Before we begin {.scrollable}

Please install and load the following packages

```{r loading}
#| echo: true

library(dplyr)
library(tidyverse)
library(scales)
library(patchwork)
library(gghighlight)
library(ggiraph)
library(ISLR2)
library(openintro)
library(janitor)
library(gapminder)
library(palmerpenguins)
```

<br> <br>

Access lecture slide from the [course landing page](https://ayushbipinpatel.github.io/GIPE-Business-Analytics/)

## About me {.scrollable}

::: columns
::: {.column width="70%"}
I am [Ayush]{.fragment fragment-index="1" style="font-size:45px"}.

[I am a researcher working at the intersection of data, law, development and economics.]{.fragment fragment-index="2" style="font-size:25px"}

[I teach Data Science using R at Gokhale Institute of Politics and Economics]{.fragment fragment-index="3" style="font-size:25px"}

[I am a [RStudio (Posit) certified tidyverse Instructor.](https://education.rstudio.com/trainers/people/patel+ayush/)]{.fragment fragment-index="4" style="font-size:25px"}

[I am a Researcher at [Oxford Poverty and Human development Initiative (OPHI)](https://ophi.org.uk/), at the University of Oxford.]{.fragment fragment-index="5" style="font-size:25px"}
:::

::: {.column width="30%"}
**Reach me**

{{< fa solid envelope >}} [ayush.ap58\@gmail.com]{style="font-size:25px"}

{{< fa solid envelope >}} [ayush.patel\@gipe.ac.in]{style="font-size:25px"}
:::
:::

## Learning Objectives {.scrollable}

::: incremental

+ Learn annotation for graphs in R
+ Learn how to combine graphs
+ Learn scaling functions in R
+ Learn how to make ggplot graphs interactive
:::

## Let's Recap{.scrollable}

+ In the data visualization lecture, you learnt how to create various types of graphs using `ggplot2`
+ Some of them include bar, graph, line graph, scatter plots etc
+ For effective data visualization and communication, any plot requires modifications
+ These include annotations on the plot, modification of axes and scales, highlighting and interactivity of the plot
+ The aim of this lecture is to move beyond making graphs, towards clear and effective visualizations

## Annotations in ggplot - Text{.scrollable}
#### [Content for this topic has been sourced from [ggplot2](https://ggplot2.tidyverse.org/reference/annotate.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ In addition to plotting your graph, you want to provide additional details to explain your graph
+ Text annotations are useful in this case
+ The `annotate()` function can be used for any kind of geometric object
+ In the `annotate()` function, the type of geom is specified first
+ Then, the positining is required (x and y coordinates in this case)
+ This is followed by the label

```{r text_annotation}
#| eval: true
#| echo: true
#| output-location: column

ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
   annotate("text", x = 4, y = 25, label = "Annotation Text")
```

## Annotations in ggplot{.scrollable}

#### [Content for this topic has been sourced from [ggplot2](https://ggplot2.tidyverse.org/reference/annotate.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Further, annotations can be customized 

```{r text_annotation_custom}
#| eval: true
#| echo: true

 ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
   annotate("text", x = 4, y = 25, label = "Annotation Text", colour = "orange", size = 8)
```

```{r text_annotation_custom1}
#| eval: true
#| echo: true

 ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
   annotate("text", x = 1:5, y = 6, label = "Annotation Text", colour = "orange", size = 3)
```

## Annotations in ggplot{.scrollable}

#### [Content for this topic has been sourced from [ggplot2](https://ggplot2.tidyverse.org/reference/annotate.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Similar to text annotation, other geoms can be used for annotations
+ However, instead of `x` and `y`, `xmin` and `xmax` is used for coordinates of the rectangle
+ Do you remember what `alpha` is used for?

```{r rect_annotation}
#| eval: true
#| echo: true
#| output-location: column

ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
   annotate("rect", xmin = 4.8, 
            xmax = 5.7,
            ymin = 10,
            ymax = 18.6, 
            alpha = .2)
```

## Annotations in ggplot{.scrollable}

#### [Content for this topic has been sourced from [ggplot2](https://ggplot2.tidyverse.org/reference/annotate.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Suppose you want to add a line segment to your graph
+ `annotate()` over here requires `x` and `xend` coordinates

```{r segment_annotation}
#| eval: true
#| echo: true
#| output-location: column

ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
   annotate("segment", x = 4.8,
            xend = 5.7,
            y = 10,
            yend = 18.6,
            colour = "red")
```

## Do it Yourself -1{.scrollable}

+ Load the `Auto` data from `ISLR2` package
+ Make a scatterplot of `horsepower` and `acceleration` 
+ To the plot, add the text `Horsepower vs. Acceleration`
+ Add a rectangle to the plot, such that it covers the area where horsepower is higher than 200 but acceleration is still lesser than 15
+ Add a line to the plot from the coordinates (50,10) to (150,20)

## Scales Functions in ggplot2 - Why?{.scrollable}

+ When you create a graph, using `ggplot2`, the axes are mapped automatically based on the data
+ However, you would often need to change the axes in order to effectively present the data
+ the `scale` functions in `ggplot2`:
  - control how the data is plotted
  - allow manipulation of axes
  - improves overall appearances of the plot for effective data communication
  
## Scales Functions in ggplot2{.scrollable}

+ Look at the scatter plot of `wt` and `mpg`

```{r scale_intro}
#| eval: true
#| echo: true

ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()
```

+ What if you want both the axes to start from 0?
+ `scale_y_continuous()` allows you to set the range for the y-axis
+ `limits` inside the `scale_y_continuous()` provides limits of the scale
+ Over here, `NA` is used to refer to the existing maximum

```{r scale_ylim}
#| eval: true
#| echo: true

ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
  scale_y_continuous(limits = c(0, NA))
```


## Scales Functions in ggplot2{.scrollable}

+ Instead of using `NA`, if you had to provide `40` as the limit for y-axis

```{r scale_ylim_specify}
#| eval: true
#| echo: true

ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
  scale_y_continuous(limits = c(0, 40))
```

## Scales Functions in ggplot2{.scrollable}

+ Setting `breaks` in the `scale_y_continuous` allows you to set what intervals the axis will have

```{r scale_ylim_breaks}
#| eval: true
#| echo: true

ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
  scale_y_continuous(breaks = seq(0, 40, 7))
```

## Scales Functions in ggplot2{.scrollable}

+ Similarly, there are other transformations for scale - reversing the scale

```{r reverses}
#| eval: true
#| echo: true

ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
  scale_y_reverse()
```

+ `scale_y_log10()` does log transformation of the scale

```{r log}
#| eval: true
#| echo: true

ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
  scale_y_log10()
```

## Scales Functions in ggplot2{.scrollable}

#### [Content for this topic has been sourced from [ggplot2](https://ggplot2.tidyverse.org/reference/scale_brewer.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ The `scale_colour_brewer()` options are useful for plotting discrete values on your graph
+ The brewer scales provide sequential colour schemes from ColorBrewer
+ Look at the two charts
+ `scale_colour_brewer` helps in effcient mapping of discrete variables

::: columns
::: {.column width="50%"}
```{r without_scale_colour_brewer}
#| eval: true
#| echo: true

ggplot(mpg, aes(x = displ, y = cty)) +
   geom_point(aes(colour = class))
```

:::

::: {.column width="50%"}

```{r scale_colour_brewer}
#| eval: true
#| echo: true

ggplot(mpg, aes(x = displ, y = cty)) +
   geom_point(aes(colour = class))+
  scale_colour_brewer()
```

:::
:::

## Do It Yourself -2{.scrollable}

+ Using the `Auto` data, plot a scatterplot between y = `weight` and x = `displacement`
+ Set the x-axis with breaks as 50
+ Set the breaks for y-axis as 5
+ What variable according to you can be used as the colour for the points? How?

## Scales Package{.scrollable}

::: columns
::: {.column width="60%"}

+ The `scales` package many scaling functions for visualizations
+ It allows for sophisticated customisation of data in a plot
+ Functions for readable and informative axes

:::

::: {.column width="40%"}

![Source : [Scale Functions for Visualizations](https://scales.r-lib.org)](https://scales.r-lib.org/logo.svg)

:::
:::

## Scales{.scrollable}
#### [Content for this topic has been sourced from [scales](https://scales.r-lib.org). Please check out the work for detailed information.]{style="font-size:15px"}

+ Look at the following chart made using `txhousing` data
+ We want to make it more readable and clear
+ The number of zeroes on the y-axis can be reduced along with the way years are represented on the x-axis

```{r scales_package_old}
#| eval: true
#| echo: true
#| output-location: column

txhousing %>% 
  mutate(date = make_date(year, month, 1)) %>% 
  group_by(city) %>% 
  filter(min(sales) > 500) %>% 
  ggplot(aes(date, sales, group = city)) + 
  geom_line(na.rm = TRUE)
```

## Scales{.scrollable}
#### [Content for this topic has been sourced from [scales](https://scales.r-lib.org). Please check out the work for detailed information.]{style="font-size:15px"}

+ Similar to the scale functions in `ggplot2`, the `scales` package has functions for `breaks` and `labels`
+ the `breaks_width` function provides a way to show every two years on the axis, while the `label_date` provides a way to show the last two digits of the year using `%y`, making it more clear
+ On the y-axis. the `cut_short_scale()` function removes the additional 0 and supplements the `K` sign

```{r scales_package_new}
#| eval: true
#| echo: true
#| output-location: column

txhousing %>% 
  mutate(date = make_date(year, month, 1)) %>% 
  group_by(city) %>% 
  filter(min(sales) > 500) %>% 
  ggplot(aes(date, sales, group = city)) + 
  geom_line(na.rm = TRUE) + 
  scale_x_date(
    NULL,
    breaks = scales::breaks_width("2 years"), 
    labels = scales::label_date("'%y")) + 
  scale_y_log10(
    "Total sales",
    labels = scales::label_number(scale_cut = scales::cut_short_scale()))
```

## Scales{.scrollable}
#### [Content for this topic has been sourced from [scales](https://scales.r-lib.org). Please check out the work for detailed information.]{style="font-size:15px"}

+ Let us try modifying another graph using `economics` data

```{r scales_economics}
#| eval: true
#| echo: true
#| output-location: column

economics %>% 
  filter(date < ymd("1970-01-01")) %>% 
  ggplot(aes(date, pce)) + 
  geom_line()
```

+ How can this be made more readable?
+ For x-axis, one option is to show the dates along with a few months, for better insights
+ For the y-axis, a label that adds the dollar sign would make the chart more readable

## Scales{.scrollable}
#### [Content for this topic has been sourced from [scales](https://scales.r-lib.org). Please check out the work for detailed information.]{style="font-size:15px"}

+ `breaks_width` sets intervals for 3 months
+ However, you might want the axes to have the date format in months
+ `label_date_short()` does the task of shortening the date lengths

```{r scales_economics_x}
#| eval: true
#| echo: true
#| output-location: column

economics %>% 
  filter(date < ymd("1970-01-01")) %>% 
  ggplot(aes(date, pce)) + 
  geom_line() + 
  scale_x_date(NULL,
    breaks = scales::breaks_width("3 months"), 
    labels = scales::label_date_short())
```

## Scales{.scrollable}
#### [Content for this topic has been sourced from [scales](https://scales.r-lib.org). Please check out the work for detailed information.]{style="font-size:15px"}

+ For the y-axis, you can set breaks as you desire using `breaks_extended()`
+ `label_dollar()` adds a dollar sign to the y-axis

```{r scales_economics_y}
#| eval: true
#| echo: true
#| output-location: column

economics %>% 
  filter(date < ymd("1970-01-01")) %>% 
  ggplot(aes(date, pce)) + 
  geom_line() + 
  scale_x_date(NULL,
    breaks = scales::breaks_width("3 months"), 
    labels = scales::label_date_short()) + 
  scale_y_continuous("Personal consumption expenditures",
    breaks = scales::breaks_extended(8),
    labels = scales::label_dollar())
```

## Do it Yourself -3{.scrollable}

+ Load the `tourism` data from `openintro`
+ Make a line graph of year and tourist spending
+ Is there any change you could make to the chart for better readability?

## Patchwork{.scrollable}

+ You have made multiple by now and want to combine them into the same graphic
+ A very easy way to do this by using `patchwork`
+ Let us learn this using our recently made plots

```{r patchwork_example}
#| eval: true
#| echo: true

p1 <- economics %>% 
  filter(date < ymd("1970-01-01")) %>% 
  ggplot(aes(date, pce)) + 
  geom_line()

p2 <- economics %>% 
  filter(date < ymd("1970-01-01")) %>% 
  ggplot(aes(date, pce)) + 
  geom_line() + 
  scale_x_date(NULL,
    breaks = scales::breaks_width("3 months"), 
    labels = scales::label_date_short())

p3 <- economics %>% 
  filter(date < ymd("1970-01-01")) %>% 
  ggplot(aes(date, pce)) + 
  geom_line() + 
  scale_x_date(NULL,
    breaks = scales::breaks_width("3 months"), 
    labels = scales::label_date_short()) + 
  scale_y_continuous("Personal consumption expenditures",
    breaks = scales::breaks_extended(8),
    labels = scales::label_dollar())

```

## Patchwork{.scrollable}

+ The usage of `patchwork` is very simple: you literally just add plots together!

```{r patchwork_1}
#| eval: true
#| echo: true

p1 + p2
```

+ You can also put the plots one below the other

```{r patchwork_2}
#| eval: true
#| echo: true

p1 / p2
```

## Patchwork{.scrollable}

+ While plots `p1` and `p2` show the intermediate steps, `p3` is the final plot
+ It would be better to have the two at the top and the final one at the bottom

```{r patchwork_3}
#| eval: true
#| echo: true
#| output-location: column

(p1 + p2) / p3
```

## Patchwork{.scrollable}

+ After combining all the plots, you would want to modify all plots at once

```{r combine_patchword}
#| eval: true
#| echo: true
#| output-location: column

patchwork <- (p1 + p2) / p3
patchwork & theme_minimal()
```

## Do it Yourself - 4{.scrollable}

+ From the `tourism` data, make line charts of `year` and `visitor_count_tho`, one for each decade
+ Combine these charts in such a way that at the top, the graph for all years is displayed and below it, there are 5 charts, one for each decade

## Highlight information - gghighlight(){.scrollable}
#### [Content for this topic has been sourced from [Hiroaki Yutani's work](https://yutannihilation.github.io/gghighlight/articles/gghighlight.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Run the following code to generate a random dataset

```{r highlight_dataset}
#| eval: true
#| echo: true

set.seed(2)
data <- purrr::map_dfr(letters, ~ data.frame(
      id = 1:500,
      value = cumsum(runif(500, -5, 5)),
      type = .,
      flag = sample(c(TRUE, FALSE), size = 500, replace = TRUE),
      stringsAsFactors = FALSE))
```

+ Suppose you want to plot the value of each id

```{r gghighlight_intro_plot}
#| eval: true
#| echo: true
#| output-location: column

ggplot(data) +
  geom_line(aes(x= id, y = value, colour = type))
```
## Highlight information - gghighlight(){.scrollable}
#### [Content for this topic has been sourced from [Hiroaki Yutani's work](https://yutannihilation.github.io/gghighlight/articles/gghighlight.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Too much clutter right?
+ You can highlight only the lines with the biggest `value` 

``` {r highlight_biggest_value}
#| eval: true
#| echo: true
#| output-location: column

data_filtered <- data %>%
  group_by(type) %>% 
  filter(max(value) > 112) %>%
  ungroup()

ggplot(data_filtered) +
  geom_line(aes(id, value, colour = type))
```

## Highlight information - gghighlight(){.scrollable}
#### [Content for this topic has been sourced from [Hiroaki Yutani's work](https://yutannihilation.github.io/gghighlight/articles/gghighlight.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ However, that takes away the context of our visualization
+ Also, very tiresome
+ `gghighlight()` comes handy here
+ Think of it as the `filter()` equivalent of `ggplot2`

```{r gghighlight}
#| eval: true
#| echo: true
#| output-location: column

ggplot(data) +
  geom_line(aes(id, value, colour = type)) +
  gghighlight(max(value) > 112)
```

+ You can customise plots as well

```{r gghighlight_customise}
#| eval: true
#| echo: true
#| output-location: column

ggplot(data) +
  geom_line(aes(id, value, colour = type)) +
  gghighlight(max(value) > 112)+
  theme_minimal()
```

## Highlight information - gghighlight(){.scrollable}
#### [Content for this topic has been sourced from [Hiroaki Yutani's work](https://yutannihilation.github.io/gghighlight/articles/gghighlight.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Similarly, you can highlight other types of geoms as well

```{r other_geom_highlight_1}
#| eval: true
#| echo: true
#| output-location: column

ggplot(mpg)+
  geom_point(aes(displ, cty))+
  gghighlight(displ >= 5)
```

+ You can use more than one condition for gghighlight

```{r other_geom_highlight_2}
#| eval: true
#| echo: true
#| output-location: column

ggplot(mpg)+
  geom_point(aes(displ, cty))+
  gghighlight(displ >= 5, cty >= 15)
```

## Highlight information - gghighlight(){.scrollable}
#### [Content for this topic has been sourced from [Hiroaki Yutani's work](https://yutannihilation.github.io/gghighlight/articles/gghighlight.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ You can select the number of values you want to highlight using `max_highlight`

```{r gghighlight_max_options}
#| eval: true
#| echo: true
#| output-location: column

ggplot(data) +
  geom_line(aes(id, value, colour = type)) +
  gghighlight(max(value), max_highlight =  5L)
```

## Do it Yourself - 5{.scrollable}

+ Use `txhousing` data from `ggplot2` and make a new dataset called `txhousing_sales` which calculates the total sales of each city for the year
+ Now, make a line chart of total sales over the years for all cities. Highlight the top 4 cities
+ From `txhousing_sales`, filter for the year 2015 and make a bar chart which shows the `total_sales` for all cities. Highlight the top 5

## Creating interactive Visualizations - ggiraph(){.scrollable}
#### [Content for this topic has been sourced from [Albert Rapp's work](https://albert-rapp.de/posts/ggplot2-tips/17_ggiraph/17_ggiraph.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Til now, we have learnt so many ways of making effective visualizations
+ When we add interactivity to our plots, it would make it very easy for people to focus on what they find important
+ `ggiraph` helps turn our graphs into an interactive visualization
+ It can be combined with many other features to connect different visualizations and communicate data-driven insights
+ Run the following code for data preperation

```{r ggiraph_data_prep}
#| eval: true
#| echo: true

gapminder_data <- gapminder::gapminder %>%
  janitor::clean_names() %>%
  mutate(id = levels(continent)[as.numeric(continent)],
    continent = forcats::fct_reorder(continent, life_exp))

mean_life_exps <- gapminder_data %>% 
  group_by(continent, year, id) %>%
  summarise(mean_life_exp = mean(life_exp)) %>%
  ungroup()
```

## Creating interactive Visualizations - ggiraph(){.scrollable}
#### [Content for this topic has been sourced from [Albert Rapp's work](https://albert-rapp.de/posts/ggplot2-tips/17_ggiraph/17_ggiraph.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Let us make a simple line chart first

``` {r simple_line_chart_ggiraph}
#| eval: true
#| echo: true
#| output-location: column

line_chart <- mean_life_exps %>%
  ggplot(aes(x = year, y = mean_life_exp, colour = continent)) +
  geom_line(linewidth = 2.5) +
  geom_point(size = 3.5) +
  theme_minimal(base_size = 18) +
  labs(x = element_blank(),
    y = 'Life expectancy (in years)',
    title = 'Life expectancy over time')
line_chart
```

## Creating interactive Visualizations - ggiraph(){.scrollable}
#### [Content for this topic has been sourced from [Albert Rapp's work](https://albert-rapp.de/posts/ggplot2-tips/17_ggiraph/17_ggiraph.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ We want to make `geom_point()` and `geom_line()` interactive
+ This can be done by 
  - changing `geom_point()` to `geom_point_interactive()`
  - changing `geom_line()` to `geom_line_interactive()`
  - adding the `data_id` aesthetic
  
``` {r simple_line_chart_ggiraph_1}
#| eval: true
#| echo: true
#| output-location: column

line_chart <- mean_life_exps %>%
  ggplot(aes(x = year, y = mean_life_exp, colour = continent, data_id = id)) +
  geom_line_interactive(linewidth = 2.5) +
  geom_point_interactive(size = 3.5) +
  theme_minimal(base_size = 18) +
  labs(x = element_blank(),
    y = 'Life expectancy (in years)',
    title = 'Life expectancy over time')
line_chart
```

``` {r girafe_object}
#| eval: true
#| echo: true
#| output-location: column

girafe(ggobj = line_chart)
```

## Creating interactive Visualizations - ggiraph(){.scrollable}
#### [Content for this topic has been sourced from [Albert Rapp's work](https://albert-rapp.de/posts/ggplot2-tips/17_ggiraph/17_ggiraph.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Currently, when you click on a line, it appears orange
+ We want to set our interactivity in such a way that when we click on a line, all other lines must fade away
+ This can be done by passing hover options to our interactive chart

```{r simple_line_chart_ggiraph_hover}
#| eval: true
#| echo: true
#| output-location: column

girafe(ggobj = line_chart,
  options = list(
    opts_hover(css = ''), ## CSS code of line we're hovering over
    opts_hover_inv(css = "opacity:0.1;"), ## CSS code of all other lines
    opts_sizing(rescale = FALSE)),
  height_svg = 6,
  width_svg = 9)

```

## Creating interactive Visualizations - ggiraph(){.scrollable}
#### [Content for this topic has been sourced from [Albert Rapp's work](https://albert-rapp.de/posts/ggplot2-tips/17_ggiraph/17_ggiraph.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Next, let us try to make an inteactive boxplot using `gapminder_data`
+ We want to make a scatterplot of life expectancy  and population across continents for the year 2007
+ First, we would make the interactive scatterplot

``` {r scatterplotplot_ggiraph1}
#| eval: true
#| echo: true
#| output-location: column

scatterplot_graph <- gapminder_data %>%
  filter(year == 2007) %>%
  ggplot(aes(x = life_exp, y = pop, fill = continent, data_id = id)) +
  geom_point_interactive(aes(col = continent), size = 3) +
  theme_minimal(base_size = 18) +
  labs(y = 'Population',
    title = 'Life expectancy vs Population in 2007')

scatterplot_graph
```

+ Next, to make all other continents' points fade away when we click on one continent points, we use hover options 

``` {r scatterplotplot_ggiraph2}
#| eval: true
#| echo: true
#| output-location: column

girafe(ggobj = scatterplot_graph,
  options = list(opts_hover(css = ''),
    opts_hover_inv(css = "opacity:0.1;"),
    opts_sizing(rescale = FALSE)),
  height_svg = 6,
  width_svg = 9)
```

## Combining `ggiraph` and `patchwork` Together{.scrollable}
#### [Content for this topic has been sourced from [Albert Rapp's work](https://albert-rapp.de/posts/ggplot2-tips/17_ggiraph/17_ggiraph.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ You can combine two interactive charts using patchwork to show the interaction of different variables at the same time
+ However, this can only be done when the underlying `data_id` is the same

```{r ggiraph_patchwork}
#| eval: true
#| echo: true

girafe(
  ggobj = scatterplot_graph + plot_spacer() + line_chart + plot_layout(widths = c(0.45, 0.1, 0.45)),
  options = list(opts_hover(css = ''),
    opts_hover_inv(css = "opacity:0.1;"), 
    opts_sizing(rescale = FALSE)),
  height_svg = 8,
  width_svg = 12)
```

## Do it Yourself - 6{.scrollable}

+ Use `penguins` data from `palmerpenguins` package. Make an interactive scatterplot of `bill_length_mm` and `body_mass_g`, using `species` for the colour
+ Next, calculate the number of species for each year. Make an interactive line chart for the same
+ Combine both the interactive charts

