---
title: "Business Analytics"
subtitle: "Writing Functions"
author: "Ayush Patel and Jayati Sharma"
date: today
date-format: " DD MMMM, YYYY"
embed-resources: true
format: 
  revealjs:
    embed-resources: true
    slide-number: c/t
    width: 1400
    theme: [serif, theme.scss]
---

## Pre-requisite {.scrollable}

::: incremental
You already....

-   Know basic and advanced data wrangling functions in R
-   Know basics of data visualization in R
-   Know univariate and multivariate linear regression
:::

## Before we begin {.scrollable}

Please install and load the following packages

```{r load}
#| echo: true

library(dplyr)
library(tidyverse)
```

<br> <br>

Access lecture slide from the [course landing page](https://ayushbipinpatel.github.io/GIPE-Business-Analytics/)

## About me {.scrollable}

::: columns
::: {.column width="70%"}
I am [Ayush]{.fragment fragment-index="1" style="font-size:45px"}.

[I am a researcher working at the intersection of data, law, development and economics.]{.fragment fragment-index="2" style="font-size:25px"}

[I teach Data Science using R at Gokhale Institute of Politics and Economics]{.fragment fragment-index="3" style="font-size:25px"}

[I am a [RStudio (Posit) certified tidyverse Instructor.](https://education.rstudio.com/trainers/people/patel+ayush/)]{.fragment fragment-index="4" style="font-size:25px"}

[I am a Researcher at [Oxford Poverty and Human development Initiative (OPHI)](https://ophi.org.uk/), at the University of Oxford.]{.fragment fragment-index="5" style="font-size:25px"}
:::

::: {.column width="30%"}
**Reach me**

{{< fa solid envelope >}} [ayush.ap58\@gmail.com]{style="font-size:25px"}

{{< fa solid envelope >}} [ayush.patel\@gipe.ac.in]{style="font-size:25px"}
:::
:::

## Learning Objectives {.scrollable}

::: incremental
-  Learn how to write functions in R
-  Learn how to perform iterations
:::

## How can we do this better?{.scrollable}

+ Assume that you are working on the following dataset 
+ You want to multiply the variables with a certain number `0.837636`
+ One way to do this is as shown below

```{r intro}
#| eval: true
#| echo: true

data <- data.frame(id = rep(letters[8 :10], each = 2), 
                   a = seq(5 ,10), b = seq(1,6), 
                   c = seq(32,37), d = seq(19,24))

new_data <- data %>%
  mutate(new_a = 0.837636 * a, 
         new_b = 0.837636 * b,
         new_c = 0.837236 * c,
         new_d = 0.837636 * d)

```

## How can we do this better?{.scrollable}

::: incremental

+ However, there are two problems with this
+ This process can get tedious as the number of operations increase
+ Also, there can be mistakes in copy pasting the number, as can be seen in the third line of `mutate()` function
+ In such cases, it is better to write functions 

:::

## Writing a basic function{.scrollable}

+ For the previous problem, we can instead write a function to perform the same operation many times
+ The idea is to give it a name of your choice, use `function()` and specify the argument `x`, then wrap up your argument and specify the operation inside `{}`
+ You will soon learn how to write functions in detail

```{r firstfunction}
#| eval: true
#| echo: true

myfirstfunction <- function(x){0.837636 * x}

 data %>%
  mutate(new_a = myfirstfunction(a), 
         new_b = myfirstfunction(b),
         new_c = myfirstfunction(c),
         new_d = myfirstfunction(d))
```

## Writing a basic function{.scrollable}

+ We can further modify our code to make it more clean

```{r firstfunctionmodified}
#| eval: true
#| echo: true

data %>%
  mutate(across(a :d, myfirstfunction))
```

## Why to write a function?{.scrollable}
#### [Content for this topic has been sourced from [Hadley Wickham's 'R for Data Science (2e)'](https://r4ds.hadley.nz/). Please check out his work for detailed information.]{style="font-size:15px"}

+ It can be given any name, which makes the code easier to understand
+ Easier to update code
+ Reduces the chances of making mistakes while copying and pasting values
+ Overall makes the code clearer for future use
+ Good rule of thumb - Consider writing a function whenever you’ve copied and pasted a block of code more than twice

## Using conditions in functions{.scrollable}

+ Let us move to writing more complex functions
+ Logic remains same; you add more conditions
+ Using `data`, you want to multiply 2 to an even number, or keep the number as it is if odd

```{r logicalfunction}
#| eval: true
#| echo: true

logic_multiply <- function(x){ifelse( x %% 2 == 0, x*2,x)}
```

+ This function can now be used to transform `data`

```{r logicalfunction_1}
#| eval: true
#| echo: true

data %>%
  mutate(across(a : d , logic_multiply))
```

## Do It Yourself -1{.scrollable}

+ Using `data`, multiply `1.77364` to all numeric variables by writing a function
+ Write a function to find square root of all nuemric variables in `data`
+ Write a function that multiplies the number by 2 if it is greater than 10, else keeps the number as it is
+ Write a function that multiplies the number by 2 if it is greater than 20, else multiplies it by 3 it is

## Data Frame Functions{.scrollable}
#### [Content for this topic has been sourced from [Hadley Wickham's 'R for Data Science (2e)'](https://r4ds.hadley.nz/). Please check out his work for detailed information.]{style="font-size:15px"}

+ Beyond vector functions, it is also possible to write data frame functions
+ Data frame functions work like vector functions: they take a data frame as the first argument, some extra arguments that say what to do with it, and return a data frame or a vector
+ Suppose you want to perform a function on `data` where you want to group the variables by `id` and get the mean of numeric variables

## Data Frame Functions{.scrollable}
#### [Content for this topic has been sourced from [Hadley Wickham's 'R for Data Science (2e)'](https://r4ds.hadley.nz/). Please check out his work for detailed information.]{style="font-size:15px"}

+ We can write the function easily using the arguments
+ Note: when we use variables from a data frame, we need to ***embrace*** them using `{{ }}`
+ This is because of **indirection**
+ Embracing a variable tells dplyr to use the value stored inside the argument, not the argument as the literal variable name

```{r dataframefunction}
#| eval: true
#| echo: true

grouped_mean <- function(dataframe, group_var, mean_var) {
  dataframe %>%
    group_by({{group_var}}) %>%
    summarize(mean({{mean_var}}))}

grouped_mean(data, id, a)
```

## Data Frame Functions{.scrollable}
#### [Content for this topic has been sourced from [Hadley Wickham's 'R for Data Science (2e)'](https://r4ds.hadley.nz/). Please check out his work for detailed information.]{style="font-size:15px"}

+ A common use for such functions is when you do your exploratory data analysis
+ You might want to see a certain set of descriptive statistics
+ Writing a function would b useful in such a case

```{r myfunction}
#| eval: true
#| echo: true

my_eda_function <- function(dataframe, variable) {
  dataframe %>%
    summarize(
      count = n(),
      minimum_value = min({{ variable}}, na.rm = TRUE),
      maximum_value = max({{ variable}}, na.rm = TRUE),
      range = max({{ variable}}, na.rm = TRUE) - min({{ variable}}, na.rm = TRUE)
    )
}

my_eda_function(data,a)
```

## Do It Yourself -2{.scrollable}

+ Using the `ChickWeight` data from `datasets`, create your own summary function that gives the minimum, maximum and range of the `weight` variable
+ Write a function that groups by `diet` and gives the `weight`

## Plot Functions{.scrollable}
#### [Content for this topic has been sourced from [Hadley Wickham's 'R for Data Science (2e)'](https://r4ds.hadley.nz/). Please check out his work for detailed information.]{style="font-size:15px"}

+ Along with returning a dataframe, functions can also return dataframes
+ Use the `diamonds` dataset from `ggplot2`
+ Suppose you want to create many histograms
+ Repeating each code with a difference in `binwidth` can be avoided

```{r plotfunctions}
#| eval: true
#| echo: false

diamonds %>% 
  ggplot(aes(x = carat)) +
  geom_histogram(binwidth = 0.1)

diamonds %>%
  ggplot(aes(x = carat)) +
  geom_histogram(binwidth = 0.05)

diamonds %>%
  ggplot(aes(x = carat)) +
  geom_histogram(binwidth = 0.3)

```

## Plot Functions - Example 1{.scrollable}
#### [Content for this topic has been sourced from [Hadley Wickham's 'R for Data Science (2e)'](https://r4ds.hadley.nz/). Please check out his work for detailed information.]{style="font-size:15px"}

+ Instead, a function an be written which reduces the task of making a plot everytime

```{r plot_function_example}
#| eval: true
#| echo: true

diamond_histogram_function <- function(data, variable, binwidth){
  data %>%
    ggplot(aes(x = {{ variable}})) +
    geom_histogram(binwidth = binwidth)
}
```

``` {r plot_show}
#| eval: true
#| echo: true

diamond_histogram_function(diamonds, carat, 0.1)
```

+ Can you guess why did we not embrace `binwidth` while writing the function?

## Plot Functions with Data Wrangling{.scrollable}
#### [Content for this topic has been sourced from [Hadley Wickham's 'R for Data Science (2e)'](https://r4ds.hadley.nz/). Please check out his work for detailed information.]{style="font-size:15px"}

+ Plot functions can be combined with data wrangling functions as well
+ To make a bar graph with the bars reordered in the descending order

```{r plot_function_example_2}
#| eval: true
#| echo: true

diamonds_sorted_bars <- function(dataframe, variable) {
  dataframe %>% 
    mutate({{ variable }} := fct_rev(fct_infreq({{ variable }})))  |>
    ggplot(aes(y = {{ variable }})) +
    geom_bar() +
    theme_minimal()
}
```

<br>

``` {r example_show}
#| eval: true
#| echo: true

diamonds %>%
  diamonds_sorted_bars(clarity)
```

## Do It Yourself -3{.scrollable}

+ Using `diamonds` dataset, write a function for making a bar graph, and then using the function, plot `cut` and `color`
+ Write a function to plot histogram of varying binwidths, and then plot variables `x` and `y` using it

## Thank You :){.center}