---
title: "Business Analytics"
subtitle: "Advanced Data Visualizations"
author: "Ayush Patel and Jayati Sharma"
date: today
date-format: " DD MMMM, YYYY"
embed-resources: true
format: 
  revealjs:
    embed-resources: true
    slide-number: c/t
    width: 1400
    theme: [serif, theme.scss]
---

## Pre-requisite {.scrollable}

::: incremental
You already....

-   Know basic and advanced data wrangling functions in R
-   Know basics of data visualization in R
-   Can write functions in R
:::

## Before we begin {.scrollable}

Please install and load the following packages

```{r loading}
#| echo: true

library(dplyr)
library(tidyverse)
library(patchwork)
library(ggiraph)
library(gghighlight)
```

<br> <br>

Access lecture slide from the [course landing page](https://ayushbipinpatel.github.io/GIPE-Business-Analytics/)

## About me {.scrollable}

::: columns
::: {.column width="70%"}
I am [Ayush]{.fragment fragment-index="1" style="font-size:45px"}.

[I am a researcher working at the intersection of data, law, development and economics.]{.fragment fragment-index="2" style="font-size:25px"}

[I teach Data Science using R at Gokhale Institute of Politics and Economics]{.fragment fragment-index="3" style="font-size:25px"}

[I am a [RStudio (Posit) certified tidyverse Instructor.](https://education.rstudio.com/trainers/people/patel+ayush/)]{.fragment fragment-index="4" style="font-size:25px"}

[I am a Researcher at [Oxford Poverty and Human development Initiative (OPHI)](https://ophi.org.uk/), at the University of Oxford.]{.fragment fragment-index="5" style="font-size:25px"}
:::

::: {.column width="30%"}
**Reach me**

{{< fa solid envelope >}} [ayush.ap58\@gmail.com]{style="font-size:25px"}

{{< fa solid envelope >}} [ayush.patel\@gipe.ac.in]{style="font-size:25px"}
:::
:::

## Learning Objectives {.scrollable}

::: incremental

+ Learn annotation for graphs in R
+ Learn how to combine graphs
+ Learn scaling functions in R
+ Learn how to make ggplot graphs interactive
:::

## Annotations in ggplot{.scrollable}

#### [Content for this topic has been sourced from [ggplot2](https://ggplot2.tidyverse.org/reference/annotate.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ In addition to plotting your graph, you want to provide additional details to explain your graph
+ Text annotations are useful in this case
+ The `annotate()` function can be used for any kind of geometric object
+ Let us begin by text annotations
+ In the `annotate()` function, the type of geom is specified first
+ Then, the positining is required (x and y coordinates in this case)
+ This is followed by the label

```{r text_annotation}
#| eval: true
#| echo: true

 ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
   annotate("text", x = 4, y = 25, label = "Annotation Text")
```

## Annotations in ggplot{.scrollable}

#### [Content for this topic has been sourced from [ggplot2](https://ggplot2.tidyverse.org/reference/annotate.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Further, annotations can be customized 

```{r text_annotation_custom}
#| eval: true
#| echo: true

 ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
   annotate("text", x = 4, y = 25, label = "Annotation Text", colour = "orange", size = 8)
```

```{r text_annotation_custom1}
#| eval: true
#| echo: true

 ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
   annotate("text", x = 1:5, y = 6, label = "Annotation Text", colour = "orange", size = 3)
```

## Annotations in ggplot{.scrollable}

#### [Content for this topic has been sourced from [ggplot2](https://ggplot2.tidyverse.org/reference/annotate.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Similar to text annotation, other geoms can be used for annotations
+ However, instead of `x` and `y`, `xmin` and `xmax` is used for coordinates of the rectangle
+ Do you remember what `alpha` is used for?

```{r rect_annotation}
#| eval: true
#| echo: true

 ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
   annotate("rect", xmin = 4.8, xmax = 5.7, ymin = 10, ymax = 18.6, alpha = .2)
```

## Annotations in ggplot{.scrollable}

#### [Content for this topic has been sourced from [ggplot2](https://ggplot2.tidyverse.org/reference/annotate.html). Please check out the work for detailed information.]{style="font-size:15px"}

+ Suppose you want to add a line segment to your graph
+ `annotate()` over here requires `x` and `xend` coordinates

```{r segment_annotation}
#| eval: true
#| echo: true

 ggplot(mtcars, aes(x = wt, y = mpg)) +
   geom_point()+
   annotate("segment", x = 4.8, xend = 5.7, y = 10, yend = 18.6, colour = "red")
```

## Scales{.scrollable}